---
title: "EDA_week4excercise"
author: "Lucia M Rodriguez Bravo"
date: "2/18/2020"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TidyFish part 1

This code is an excercise for the R class taught at UABC on Spring 2020. It is meant to tidy a database.

# Data extraction
First we're calling the package to be used
```{r}
library(tidyverse)
library(dbplyr)
```

Second, calling up the dataset from a csv file 
```{r}
MessyData<-read.csv("week4/data/CompDec2011_2015.csv",na.strings = "missing")
#Note that on some sets NA values fill empty spaces, while others have "missing" written. The Addition of na.strings="missing" corrects this
```

The data consists of frequency of fish observed in the month of December per site per year.

# Data transformation
The data is in wide format. Next, it will be transformed to long format
```{r}
TidyData<-pivot_longer(MessyData, -c("DISTINCT.SPP.LIST","Month"), names_to=c("Site","Year"), names_sep="\\.", values_to="Quantity",values_drop_na = TRUE)
```

Notes on this code: 
Originally the code was written as follows:
```{r}
#TidyData<-pivot_longer(MessyData, X119.2011:X89.2013, names_to=c("Site","Year"), names_sep="\\.", values_to="Quantity")
```
To include rather than exclude data (Xthis:Xthat versus -c("A","B) ) resulted in too many columns.
The addition of values_drop_na=TRUE eliminates unnecessary rows by not including NAN values.

# Data processing
The next task is to only have data from 2011
```{r}
TidyData2011<-filter(TidyData,Year==2011)
```
Since the Month column is redundant, we'll get rid of it
```{r}
TidyData2011<-select(TidyData2011,Species=DISTINCT.SPP.LIST,Year,Site,Quantity)
```
Note also that I chose to change the name of the species column to something I liked better :)

A little fact checking:
```{r}
str(TidyData2011)
unique(TidyData2011$Year)
```
