---
title: "EDA_week4excercise2"
author: "Lucia M Rodriguez Bravo"
date: "2/18/2020"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TidyFish part 2

## Data extraction
First we're calling the packages to be used
```{r}
library(tidyverse)
library(dbplyr)
```

Second, calling up the datasets from a csv file 
```{r}
A<-read.csv("week4.2/data/FishAb.csv")
B<-read.csv("week4.2/data/FishAbtrscts.csv")
C<-read.csv("week4.2/data/SppCodes.csv")
```
Exploring:
```{r}
str(A)
str(B)
str(C)
```

## Tidying the datasets
I will need to:

1.extract fish smaller than 20cm (dataset A)
2. add them by date, site and species (date+site B, species A)

Dataset C has no relevant information.

  For dataset A, we need to filter fish bigger than 20cm.
Since total length (TL) is no longer relevant, it can be done away with.
```{r}
A<-filter(A,TL<20)
A<-select(A,Transect,ComName,Freq)

```
  For dataset B, we need to reduce it to the variables we care about. 
Since both datasets share transect number as a variable, the name gets standardized on this step too
```{r}
B<-select(B,Transect=TransectNum,Date,Site)
```
  After both datasets got cleaned up, they can be joined by their shared variable (Transect), which can then be done away with
```{r}
Fish<-left_join(A,B,by="Transect")
Fish<-select(Fish,-Transect)
```

## Adding fish counts
  Our single dataset now has all relevant information, but fish counts are still broken up, so we add them
```{r}
Fish<-aggregate(Fish$Freq,by=list(Fish$ComName,Fish$Date,Fish$Site),sum)
Fish<-rename(Fish,ComName=Group.1,Date=Group.2,Site=Group.3,Freq=x)
```
Note: aggregate renamed my columns, which I undid by re-renaming them rather than figuring out why. It works :) I am also pretty sure this isn't how I was supposed to do this but I did and I'm reasonably sure it's correct.
```{r}
str(Fish)
```
